### 2024-08-16 supplemental_material.Rmd

---
title: "Supplemental material"
author: "Courtney Goodridge"
date: "16/08/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Supplemental material

This analysis script contains analysis and plotting for the supplemental material associated with the following manuscript:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.