### 2024-08-30 supplemental_material.Rmd
new code for plotting and model fitting with truncation

---
title: "Supplemental material"
author: "Courtney Goodridge"
date: "16/08/2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Supplemental material

This analysis script contains analysis and plotting for the supplemental material associated with the following manuscript:

## Packages

```{r}
if(!require(here)) install.packages("here")
library(here)

if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if(!require(dplyr)) install.packages("dplyr")
library(dplyr)

if(!require(tidyr)) install.packages("tidyr")
library(tidyr)

if(!require(viridis)) install.packages("viridis")
library(viridis)

if(!require(gridExtra)) install.packages("gridExtra")
library(gridExtra)

if(!require(readr)) install.packages("readr")
library(readr)

if(!require(plyr)) install.packages("plyr")
library(plyr)

if(!require(stringr)) install.packages("stringr")
library(stringr)

if(!require(readxl)) install.packages("readxl")
library(readxl)

if(!require(data.table)) install.packages("data.table")
library(data.table)

if(!require(xfun)) install.packages("xfun")
library(xfun)

if(!require(cowplot)) install.packages("cowplot")
library(cowplot)

if(!require(gganimate)) install.packages("gganimate")
library(gganimate)

if(!require(ggExtra)) install.packages("ggExtra")
library(ggExtra)

if(!require(plotly)) install.packages("plotly")
library(plotly)

if(!require(brms)) install.packages("brms")
library(brms)

if(!require(bayestestR)) install.packages("bayestestR")
library(bayestestR)

if(!require(emmeans)) install.packages("emmeans")
library(emmeans)

if(!require(ROSE)) install.packages("ROSE")
library(ROSE)

if(!require(Rmisc)) install.packages("Rmisc")
library(Rmisc)

if(!require(metafolio)) install.packages("metafolio")
library(metafolio)

if(!require(tibble)) install.packages("tibble")
library(tibble)

if(!require(modelr)) install.packages("modelr")
library(modelr)

if(!require(tidybayes)) install.packages("tidybayes")
library(tidybayes)

if(!require(mgcv)) install.packages("mgcv")
library(mgcv)

if(!require(ciTools)) install.packages("ciTools")
library(ciTools)

if(!require(gratia)) install.packages("gratia")
library(gratia)

if(!require(scales)) install.packages("scales")
library(scales)

if(!require(patchwork)) install.packages("patchwork")
library(patchwork)

if(!require(ggforce)) install.packages("ggforce")
library(ggforce)
```

# Load vehicle data

This provides timing responses to takeovers and collision data. We also remove participants that were removed in the gaze entropy analysis.

```{r}
# response data frame for vehicle and eye data per trial
response.vehicle.eye <- fread(file = here::here("takeover_performance_cogload_time_budget/Data/takeover.responses.critical.lat.long.csv"))

# removing participants from vehicle and gaze data
response.vehicle.eye <- response.vehicle.eye %>%
  dplyr::filter(ppid != 34, ppid != 39, ppid != 47) %>%
  tidyr::drop_na()

# make collision data 1s and 0s
response.vehicle.eye <- response.vehicle.eye %>%
  dplyr::mutate(collision = case_when(collision == TRUE ~ 1,
                                      collision == FALSE ~ 0))

# computing outliers for braking reaction time
response.vehicle.eye <-  response.vehicle.eye %>%
  #dplyr::filter(collision == 0) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(n_back, ttc_criticality.x) %>%
  dplyr::mutate(mean.rt = mean(rt), sd.rt = sd(rt)) %>%
  dplyr::mutate(sd.rt.4 =  sd.rt * 4) %>%
  dplyr::mutate(mean_4_sd = mean.rt + sd.rt.4) %>%
  dplyr::mutate(rt.outlier = rt > mean_4_sd)

# calculating outliers for hands on reaction time
response.vehicle.eye <-  response.vehicle.eye %>%
  #dplyr::filter(collision == 0) %>%
  tidyr::drop_na() %>%
  dplyr::group_by(n_back, ttc_criticality.x) %>%
  dplyr::mutate(mean.ho = mean(ho.rt), sd.ho = sd(ho.rt)) %>%
  dplyr::mutate(sd.ho.4 =  sd.ho * 4) %>%
  dplyr::mutate(mean_4_sd = mean.ho + sd.ho.4) %>%
  dplyr::mutate(ho.outlier = ho.rt > mean_4_sd)

# removing outliers
response.vehicle.eye <- response.vehicle.eye %>%
  dplyr::filter(rt.outlier == FALSE, ho.outlier == FALSE, ho.rt != 0, collision == 0, ho.rt > .15) %>%
  dplyr::mutate(ttc_criticality_factor = as.factor(ttc_criticality.x))
```

## Braking reaction time

```{r}
# loading model
braking_rt_shifted_lognormal_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/braking_rt_shifted_lognormal_maximal.rds"))

# posterior predictive checks
brt_stat_grouped_ttc <- pp_check(braking_rt_shifted_lognormal_maximal, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
brt_stat_grouped_nback <- pp_check(braking_rt_shifted_lognormal_maximal, type = "stat_grouped", stat = "mean", group = "n_back")

# predicted global mean RTs implied by the model # setting re_formula = NA tells the function to not use any random effects - here we plot the TTC specific distributions 
predicted_braking_rt_ttc <- braking_rt_shifted_lognormal_maximal %>% 
  epred_draws(newdata = expand_grid(ttc_criticality_factor = c(as.factor(3), as.factor(5)),
                                    n_back = FALSE),
              re_formula = NA)

predicted_braking_rt_ttc %>% mean_hdi()

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR MEAN BRT

# mean BRT for critcality
sm_fig_1_1 <- ggplot(brt_stat_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(mean_brt = mean(rt)), mapping = aes(xintercept = mean_brt, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(1.0, 1.5), breaks = seq(1.0, 1.4, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(RT[B]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.585, .70), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean BRT for N-back
sm_fig_1_2 <- ggplot(brt_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(mean_brt = mean(rt)), mapping = aes(xintercept = mean_brt, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(1.0, 1.5), breaks = seq(1.0, 1.4, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(RT[B]~ "(s)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig1 <- sm_fig_1_1 / sm_fig_1_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig1.tiff"), plot = sm_fig1, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS

pp_check(braking_rt_shifted_lognormal_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(braking_rt_shifted_lognormal_maximal, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_brt <- response.vehicle.eye %>% 
  select(rt, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(braking_rt_shifted_lognormal_maximal, ndraws = 100)

sm_fig2_1 <- ggplot() + 
  geom_density(yrep_brt, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(rt, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(RT[B]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.94, .71), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig2_2 <- ggplot() + 
  geom_density(yrep_brt, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(rt, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(RT[B]~ "(s)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig2 <- sm_fig2_1 / sm_fig2_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig2.tiff"), plot = sm_fig2, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS

# - change NA to observed data - then use legend to identify y versus y_rep
brt_hist <- pp_check(braking_rt_shifted_lognormal_maximal, type = "hist", ndraws = 11)

brt_hist_df <- brt_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig3 <- ggplot(brt_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(RT[B]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.945, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig3.tiff"), plot = sm_fig3, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## Hands on reatcion time

```{r}
# loading model
hands_on_rt_model <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/handson_rt_shifted_lognormal_maximal.rds"))

# posterior predictive checks
hort_stat_grouped_ttc <- pp_check(hands_on_rt_model, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
hort_stat_grouped_nback <- pp_check(hands_on_rt_model, type = "stat_grouped", stat = "mean", group = "n_back")

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR MEAN BRT

# mean HORT for critcality
sm_fig_4_1 <- ggplot(hort_stat_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(mean_hort = mean(ho.rt)), mapping = aes(xintercept = mean_hort, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(1.1, 1.8), breaks = seq(1.1, 1.8, .2), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(RT[HO]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.593, .65), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean BRT for N-back
sm_fig_4_2 <- ggplot(hort_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(mean_hort = mean(ho.rt)), mapping = aes(xintercept = mean_hort, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(1.1, 1.8), breaks = seq(1.1, 1.8, .2), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(RT[HO]~ "(s)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig4 <- sm_fig_4_1 / sm_fig_4_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig4.tiff"), plot = sm_fig4, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS

pp_check(hands_on_rt_model, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(hands_on_rt_model, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_hort <- response.vehicle.eye %>% 
  select(ho.rt, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(hands_on_rt_model, ndraws = 100)

sm_fig5_1 <- ggplot() + 
  geom_density(yrep_hort, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(rt, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  xlim(0, 8) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(RT[HO]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.94, .69), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig5_2 <- ggplot() + 
  geom_density(yrep_hort, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(rt, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  xlim(0, 8) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(RT[HO]~ "(s)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig5 <- sm_fig5_1 / sm_fig5_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig5.tiff"), plot = sm_fig5, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS

# - change NA to observed data - then use legend to identify y versus y_rep
hort_hist <- pp_check(hands_on_rt_model, type = "hist", ndraws = 11)

hort_hist_df <- hort_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig6 <- ggplot(hort_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  xlim(0, 8) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(RT[HO]~ "(s)")) +
  theme_bw() +
  theme(legend.position = c(.945, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig6.tiff"), plot = sm_fig6, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## Mean deceleration

```{r}
# loading model
mean_decel_model <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/mean_decel_skewnormal_reg_priors_maximal.rds"))

# posterior predictive checks
decel_stat_grouped_ttc <- pp_check(mean_decel_model, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
decel_stat_grouped_nback <- pp_check(mean_decel_model, type = "stat_grouped", stat = "mean", group = "n_back")

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR MEAN BRT

# mean decel for critcality
sm_fig_7_1 <- ggplot(decel_stat_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(mean_decel = mean(mean_decel)), mapping = aes(xintercept = mean_decel, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(-4, -1), breaks = seq(-1, -4, -1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(M[D] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.593, .65), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean BRT for N-back
sm_fig_7_2 <- ggplot(decel_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(mean_decel = mean(mean_decel)), mapping = aes(xintercept = mean_decel, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(-3, -1.75), breaks = seq(-3, -2, .25), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(M[D] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig7 <- sm_fig_7_1 / sm_fig_7_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig7.tiff"), plot = sm_fig7, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS

pp_check(mean_decel_model, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(mean_decel_model, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_decel <- response.vehicle.eye %>% 
  select(mean_decel, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(mean_decel_model, ndraws = 100)

sm_fig8_1 <- ggplot() + 
  geom_density(yrep_decel, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(mean_decel, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  #xlim(-9, 2) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(M[D] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.94, .69), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig8_2 <- ggplot() + 
  geom_density(yrep_decel, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(mean_decel, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  #xlim(-9, 2) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(M[D] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig8 <- sm_fig8_1 / sm_fig8_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig8.tiff"), plot = sm_fig8, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS

# - change NA to observed data - then use legend to identify y versus y_rep
decel_hist <- pp_check(mean_decel_model, type = "hist", ndraws = 11)

decel_hist_df <- decel_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig9 <- ggplot(decel_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  #xlim(-8.5, 1) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(M[D] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.815, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig9.tiff"), plot = sm_fig9, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## Mean lateral accelerations

```{r}
# loading model
lat_model <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/mean_lat_acc_lognormal_maximal.rds"))

# posterior predictive checks
lat_stat_grouped_ttc <- pp_check(lat_model, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
lat_stat_grouped_nback <- pp_check(lat_model, type = "stat_grouped", stat = "mean", group = "n_back")

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR MEAN lat

# mean lateral for criticality
sm_fig_10_1 <- ggplot(lat_stat_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(lat = mean(mean_lat_acc)), mapping = aes(xintercept = lat, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(.1, .4), breaks = seq(.1, .4, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(M[LA] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.593, .65), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean lateral for N-back
sm_fig_10_2 <- ggplot(lat_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(lat = mean(mean_lat_acc)), mapping = aes(xintercept = lat, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(.1, .4), breaks = seq(.1, .4, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(M[LA] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig10 <- sm_fig_10_1 / sm_fig_10_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig10.tiff"), plot = sm_fig10, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS

pp_check(lat_model, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(lat_model, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_lat <- response.vehicle.eye %>% 
  select(mean_lat_acc, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(lat_model, ndraws = 100)

sm_fig11_1 <- ggplot() + 
  geom_density(yrep_lat, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(mean_lat_acc, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  xlim(0, 3) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(M[LA] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.94, .69), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig11_2 <- ggplot() + 
  geom_density(yrep_lat, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(mean_lat_acc, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  xlim(0, 3) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(M[LA] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig11 <- sm_fig11_1 / sm_fig11_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig11.tiff"), plot = sm_fig11, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS

# - change NA to observed data - then use legend to identify y versus y_rep
lat_hist <- pp_check(lat_model, type = "hist", ndraws = 11)

lat_hist_df <- lat_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig12 <- ggplot(lat_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  scale_x_continuous(limits = c(0, 2), breaks = seq(0, 1.5, .5), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(M[LA] ~"("~m/s^2~")")) +
  theme_bw() +
  theme(legend.position = c(.945, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

  # plot saving
  ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig12.tiff"), plot = sm_fig12, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## SDLP

```{r}
# loading model
sdlp_model <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/sdlp_lognormal_maximal.rds"))

# posterior predictive checks
sdlp_stat_grouped_ttc <- pp_check(sdlp_model, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
sdlp_stat_grouped_nback <- pp_check(sdlp_model, type = "stat_grouped", stat = "mean", group = "n_back")

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR MEAN SDLP

# mean SDLP for criticality
sm_fig_13_1 <- ggplot(sdlp_stat_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(sdlp = mean(sdlp)), mapping = aes(xintercept = sdlp, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(.2, .6), breaks = seq(.2, .6, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(SD[LP]~ "(m)")) +
  theme_bw() +
  theme(legend.position = c(.92, .70), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean SDLP for N-back
sm_fig_13_2 <- ggplot(sdlp_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(sdlp = mean(sdlp)), mapping = aes(xintercept = sdlp, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(.2, .6), breaks = seq(.2, .6, .1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(SD[LP]~ "(m)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig13 <- sm_fig_13_1 / sm_fig_13_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig13.tiff"), plot = sm_fig13, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS

pp_check(sdlp_model, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(sdlp_model, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_sdlp <- response.vehicle.eye %>% 
  select(sdlp, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(sdlp_model, ndraws = 100)

sm_fig14_1 <- ggplot() + 
  geom_density(yrep_sdlp, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sdlp, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(SD[LP]~ "(m)")) +
  theme_bw() +
  theme(legend.position = c(.94, .72), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig14_2 <- ggplot() + 
  geom_density(yrep_sdlp, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sdlp, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(SD[LP]~ "(m)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig14 <- sm_fig14_1 / sm_fig14_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig14.tiff"), plot = sm_fig14, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS

sdlp_hist <- pp_check(sdlp_model, type = "hist", ndraws = 11)

sdlp_hist_df <- sdlp_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig15 <- ggplot(sdlp_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  xlim(0, 5) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[LP]~ "(m)")) +
  theme_bw() +
  theme(legend.position = c(.945, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig15.tiff"), plot = sm_fig15, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## SD of horizontal gaze - pre-takeover

```{r}
sd_yaw_automation_lognormal_maximal <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_automation_lognormal_maximal.rds"))

# posterior predictive checks
yaw_pre_stat_grouped_nback <- pp_check(sd_yaw_automation_lognormal_maximal, type = "stat_grouped", stat = "mean", group = "n_back")

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean SSD of yaw for N-back
sm_fig16 <- ggplot(yaw_pre_stat_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(sd_yaw_automation = mean(sd_yaw_automation)), mapping = aes(xintercept = sd_yaw_automation, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(3, 12), breaks = seq(4, 12, 2), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig16.tiff"), plot = sm_fig16, width = 16, height = 6, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS 
pp_check(sd_yaw_automation_lognormal_maximal, type = "dens_overlay_grouped", group = "n_back", ndraws = 100)

yrep_yaw_pre <- response.vehicle.eye %>% 
  select(sd_yaw_automation, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(sd_yaw_automation_lognormal_maximal, ndraws = 100)

sm_fig17 <- ggplot() + 
  geom_density(yrep_yaw_pre, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sd_yaw_automation, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig17.tiff"), plot = sm_fig17, width = 16, height = 6, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS
yaw_pre_hist <- pp_check(sd_yaw_automation_lognormal_maximal, type = "hist", ndraws = 11)

yaw_pre_hist_df <- yaw_pre_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig18 <- ggplot(yaw_pre_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  xlim(-5, 40) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.945, .905), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig18.tiff"), plot = sm_fig18, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## SD of horizontal gaze - post-takeover (Gaussian)

```{r}
# loading model
yaw_post_model <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_manual_gaussian_maximal.rds"))

# posterior predictive checks
yaw_post_grouped_ttc <- pp_check(yaw_post_model, type = "stat_grouped", stat = "mean", group = "ttc_criticality_factor")
yaw_post_grouped_nback <- pp_check(yaw_post_model, type = "stat_grouped", stat = "mean", group = "n_back")

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

### STAT GROUPED PLOTS FOR sd of yaw
sm_fig_19_1 <- ggplot(yaw_post_grouped_ttc[["data"]] %>%
         dplyr::mutate(ttc_criticality_factor = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(ttc_criticality_factor) %>%
               dplyr::summarise(sd_yaw = mean(sd_yaw_manual)), mapping = aes(xintercept = sd_yaw, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(8, 12), breaks = seq(8, 12, 1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.92, .70), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

# mean lateral for N-back
sm_fig_19_2 <- ggplot(yaw_post_grouped_nback[["data"]] %>%
         dplyr::mutate(n_back = as.factor(group)), mapping = aes(value)) +
  geom_histogram(aes(fill = "T(y_rep)"),col = "black") + 
  geom_vline(response.vehicle.eye %>% 
               dplyr::group_by(n_back) %>%
               dplyr::summarise(sd_yaw = mean(sd_yaw_manual)), mapping = aes(xintercept = sd_yaw, fill = "T(y)"), col = "darkblue", size = 2) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  scale_fill_manual(name = "T = mean", values = c("darkblue", "lightblue"), labels = c(expression(T(y)), expression(T(y[rep])))) +
  scale_x_continuous(limits = c(8, 12), breaks = seq(8, 12, 1), labels = label_number(accuracy = 0.1)) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig19 <- sm_fig_19_1 / sm_fig_19_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig19.tiff"), plot = sm_fig19, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### DENSITY POSTERIOR PLOTS 
pp_check(yaw_post_model, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(yaw_post_model, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_yaw_post <- response.vehicle.eye %>% 
  select(sd_yaw_manual, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(yaw_post_model, ndraws = 100)

sm_fig20_1 <- ggplot() + 
  geom_density(yrep_yaw_post, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sd_yaw_manual, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  #xlim(-5, 25) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.94, .69), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig20_2 <- ggplot() + 
  geom_density(yrep_yaw_post, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sd_yaw_manual, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  #xlim(-5, 25) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig20 <- sm_fig20_1 / sm_fig20_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig20.tiff"), plot = sm_fig20, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS
lat_hist <- pp_check(yaw_post_model, type = "hist", ndraws = 11)

lat_hist_df <- lat_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig21 <- ggplot(lat_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  #xlim(-5, 30) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.611, .955), legend.direction = "horizontal", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig21.tiff"), plot = sm_fig21, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```

## SD of horizontal gaze - post-takeover (Truncated Gaussian)

```{r}
# loading model
yaw_post_model_trunc <- readRDS(here::here("takeover_performance_cogload_time_budget/Models/sd_yaw_manual_trunc_gaussian_maximal.rds"))

ttc_labs <- c("TTC = 3 s", "TTC = 5 s")
names(ttc_labs) <- c(as.factor(3), as.factor(5))

n_back_labs <- c("No N-back", "N-back")
names(n_back_labs) <- c(FALSE, TRUE)

### DENSITY POSTERIOR PLOTS 
pp_check(yaw_post_model_trunc, type = "dens_overlay_grouped", group = "n_back", ndraws = 100) 
pp_check(yaw_post_model_trunc, type = "dens_overlay_grouped", group = "ttc_criticality_factor", ndraws = 100)

yrep_yaw_post <- response.vehicle.eye %>% 
  select(sd_yaw_manual, n_back, ttc_criticality_factor, ppid) %>% 
  add_predicted_draws(yaw_post_model_trunc, ndraws = 100)

sm_fig22_1 <- ggplot() + 
  geom_density(yrep_yaw_post, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sd_yaw_manual, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ ttc_criticality_factor, labeller = labeller(ttc_criticality_factor = ttc_labs)) +
  #xlim(-5, 25) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("A") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.94, .69), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig22_2 <- ggplot() + 
  geom_density(yrep_yaw_post, mapping = aes(.prediction, group = .draw, col = "y_rep"), alpha = .8) +
  geom_density(response.vehicle.eye, mapping = aes(sd_yaw_manual, col = "y"), size = 1) +
  scale_colour_manual(name = " ", values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ n_back, labeller = labeller(n_back = n_back_labs)) +
  #xlim(-5, 25) +
  scale_y_continuous(expand = c(0, 0)) +
  ggtitle("B") +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = "none", axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

sm_fig22 <- sm_fig22_1 / sm_fig22_2 + plot_layout(axes = "collect")

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig22.tiff"), plot = sm_fig22, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')

### SIMULATED NEW PARTICIPANTS
lat_hist <- pp_check(yaw_post_model_trunc, type = "hist", ndraws = 11)

lat_hist_df <- lat_hist[["data"]] %>%
  mutate(across(everything(), ~ replace_na(., 12)))

sim_labs <- c(seq(1, 11, 1), "Observed data")
names(sim_labs) <- c(seq(1, 11, 1), 12)

sm_fig23 <- ggplot(lat_hist_df, mapping = aes(value, fill = is_y_label)) +
  geom_histogram(col = "black") +
  scale_fill_manual(name = NULL, values = c("darkblue", "lightblue"), labels = c(expression(y), expression(y[rep]))) +
  facet_wrap(~ rep_id, labeller = labeller(rep_id = sim_labs)) +
  #xlim(-5, 30) +
  scale_y_continuous(expand = c(0, 0)) +
  ylab(" ") +
  xlab(expression(SD[YAW]~ "(°)")) +
  theme_bw() +
  theme(legend.position = c(.945, .565), axis.title.x = element_text(size = 11), axis.text.x = element_text(size = 11), axis.text.y = element_blank(), axis.ticks.y = element_blank(),axis.ticks = element_blank(), title = element_text(size = 18), legend.title = element_text(size = 12), legend.text = element_text(size = 12), legend.key = element_blank(), legend.key.width = unit(0.3, 'cm'), legend.key.size = unit(0.1, 'cm'), plot.title = element_text(size = 15, face = "bold"), strip.text = element_text(face = "bold", size = 10))

# plot saving
ggsave(here::here("takeover_performance_cogload_time_budget/Supplemental material/Figures/sm_fig23.tiff"), plot = sm_fig23, width = 16, height = 12, units = 'cm', dpi = 300, type = 'cairo')
```